\textsc{Taking Dyads Seriously}

### 2018-08-29 prezzy/netsMatter.tex
preeze deck

% tables
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{multirow}

@@

% references to graphics
\makeatletter
\def\input@path{{/Users/janus829/Dropbox/Research/netModels/summResults/}, {/Users/s7m/Dropbox/Research/netModels/summResults/}, {/Users/mdw/Dropbox/Research/netModels/summResults/}}
\graphicspath{{/Users/janus829/Dropbox/Research/netModels/summResults/}, {/Users/s7m/Dropbox/Research/netModels/summResults/},{/Users/mdw/Dropbox/Research/netModels/summResults/}}

\title[AMEN]{\textsc{Taking Dyads Seriously}}
\author[Minhas et al.]{Shahryar Minhas$^\dag$, and all the others \\ Duke University \\ $^\dag$ Department of Political Science \\ $^\ddag$ Department of Statistical Science}

@@

\begin{itemize}
\item - biased effects estimation
\item - uncalibrated confidence intervals
\item - poor predictive performance
\item - inaccurate description of network phenomena
\end{itemize}

@@

\begin{tabular}{lll}
Thompson \& Walker (1982) & Beck et al. (1998) & Snijders (2011) \\
Frank \& Strauss (1986) & Signorino (1999) & Erikson et al. (2014) \\
Kenny (1996) & Li \& Loken (2002) & Aronow et al. (2015) \\
Krackhardt (1998) & Hoﬀ \& Ward (2004) & Athey et al. (2016) \\
\end{tabular}

@@

\begin{itemize}
  \item - Nodal and dyadic dependencies in networks
  \item \qquad -  Can model using the ``A'' in AME 
  \item - Third order dependencies
  \item \qquad - Can model using the ``M'' in AME
  \item - Application
\end{itemize}

@@

We use this model to form the additive effects portion of AME

@@

Each node $i$ has an unknown latent factor

@@

\vspace{-5mm}
The probability of a tie from $i$ to $j$ depends on their latent factors

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Swiss Climate Change Application}

Cross-sectional network measuring whether an actor indicated that they collaborated with another during the policy design of the Swiss CO$_{2}$ act (Ingold 2008)

\begin{figure}[ht]
  \centering
  \begin{tabular}{cc}
  \includegraphics[width=.47\textwidth]{dvNet_outDegree} & 
  \includegraphics[width=.44\textwidth]{dvNet_inDegree}
  \end{tabular}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Parameter Estimates}

% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Sun Aug 21 03:32:43 2016
\vspace{-15mm}
\begin{table}[ht]
\centering
\begingroup\scriptsize
\begin{tabular}{lcccccc}
   & Expected & \multirow{2}{*}{Logit} & \multirow{2}{*}{MRQAP} & \multirow{2}{*}{LSM} & \multirow{2}{*}{ERGM} & \multirow{2}{*}{AME} \\ 
   & Effect & & & & & \\
  \hline
  \textbf{Conflicting policy preferences} &  &  &  &  &  &  \\ 
  $\;\;\;\;$ Business vs. NGO & \color{red}{$-$} & -0.86 & \color{red}{-0.87$^{\ast}$} & \color{red}{-1.37$^{\ast}$} & \color{red}{-1.11$^{\ast}$} & \color{red}{-1.37$^{\ast}$} \\ 
  $\;\;\;\;$ Opposition/alliance & \color{blue}{$+$} &  \color{blue}{1.21$^{\ast}$} & \color{blue}{1.14$^{\ast}$} & 0.00 & \color{blue}{1.22$^{\ast}$} & \color{blue}{1.08$^{\ast}$} \\ 
  $\;\;\;\;$ Preference dissimilarity & \color{red}{$-$} & -0.07 & -0.60 & \color{red}{-1.76$^{\ast}$} & -0.44 & \color{red}{-0.79$^{\ast}$} \\ 
  \textbf{Transaction costs} &  &  &  &  &  &  \\ 
  $\;\;\;\;$ Joint forum participation & \color{blue}{$+$} & \color{blue}{0.88$^{\ast}$} & \color{blue}{0.75$^{\ast}$} & \color{blue}{1.51$^{\ast}$} & \color{blue}{0.90$^{\ast}$} & \color{blue}{0.92$^{\ast}$} \\ 
  \textbf{Influence} & &  &  &  &  &  \\ 
  $\;\;\;\;$ Influence attribution & \color{blue}{$+$} & \color{blue}{1.20$^{\ast}$} & \color{blue}{1.29$^{\ast}$} & 0.08 & \color{blue}{1.00$^{\ast}$} & \color{blue}{1.09$^{\ast}$} \\ 
  $\;\;\;\;$ Alter's influence indegree & \color{blue}{$+$} & \color{blue}{0.10$^{\ast}$} & \color{blue}{0.11$^{\ast}$} & 0.01 & \color{blue}{0.21$^{\ast}$} & \color{blue}{0.11$^{\ast}$} \\ 
  $\;\;\;\;$ Influence absolute diff. & \color{red}{$-$} & \color{red}{-0.03$^{\ast}$} & \color{red}{-0.06$^{\ast}$} & 0.04 & \color{red}{-0.05$^{\ast}$} & \color{red}{-0.07$^{\ast}$} \\ 
  $\;\;\;\;$ Alter = Government actor & \color{blue}{$+$} & \color{blue}{0.63$^{\ast}$} & 0.68 & -0.46 & \color{blue}{1.04$^{\ast}$} & 0.55 \\ 
  \textbf{Functional requirements} &  &  &  &  &  & \\ 
  $\;\;\;\;$ Ego = Environmental NGO & \color{blue}{$+$} & \color{blue}{0.88$^{\ast}$} & 0.99 & -0.60 & \color{blue}{0.79$^{\ast}$} & 0.67 \\ 
  $\;\;\;\;$ Same actor type & \color{blue}{$+$} & \color{blue}{0.74$^{\ast}$} & \color{blue}{1.12$^{\ast}$} & \color{blue}{1.17$^{\ast}$} & \color{blue}{0.99$^{\ast}$} & \color{blue}{1.04$^{\ast}$} \\ 
   \hline
\end{tabular}
\endgroup
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Latent Factor Visualization}

  \centering
  \vspace{-10mm}
  \includegraphics[width=.7\textwidth]{ameFitSR_2_UV}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Out of Sample Performance Assessment}

\begin{itemize}
  \item - Randomly divide the $n \times (n-1)$ data points into $S$ sets of roughly equal size, letting $s_{ij}$ be the set to which pair $\{ij\}$ is assigned.
  \item - For each $s \in \{1, \ldots, S\}$:
  \begin{itemize}
    \item - Obtain estimates of the model parameters conditional on $\{y_{ij} : s_{ij} \neq s\}$, the data on pairs not in set $s$.
    \item - For pairs $\{kl\}$ in set $s$, let $\hat y_{kl} = E[y_{kl} | \{y_{ij} : s_{ij} \neq s\}]$, the predicted value of $y_{kl}$ obtained using data not in set $s$.
  \end{itemize}
\end{itemize}

This procedure generates a sociomatrix of out-of-sample predictions of the observed data

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Performance Comparison}

\begin{tabular}{cc}
\includegraphics[width=.5\textwidth]{roc_outSample} & 
\includegraphics[width=.5\textwidth]{rocPr_outSample} 
\end{tabular}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Network Dependencies}

  \centering
  \includegraphics[width=1\textwidth]{netPerfCoef}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{What's Next?}

  \begin{itemize}
    \item - Extending AMEN to handle changing actor compositions
    \item - Generalize multiplicative effects estimation over tensors
  \end{itemize}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Standard Network Dependence Measures}
\vspace{-5mm}
\includegraphics[width=1\textwidth]{ggGofAll_preeze}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Simulation Comparison}

\includegraphics[width=1\textwidth]{ameVergmSim.png}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{AMEN v LSM Performance}

  \begin{tabular}{cc}
  \includegraphics[width=.5\textwidth]{roc_latSpace_outSample} & 
  \includegraphics[width=.5\textwidth]{rocPr_latSpace_outSample}
  \end{tabular}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{AMEN versus LSM Net Dependence}

  \includegraphics[width=1\textwidth]{netPerfCoef_latSpace}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{AMEN varying $K$ Performance}

  \begin{tabular}{cc}
  \includegraphics[width=.5\textwidth]{roc_ameSR_outSample} & 
  \includegraphics[width=.5\textwidth]{rocPr_ameSR_outSample}
  \end{tabular}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{AMEN varying $K$ Net Dependence}

  \includegraphics[width=1\textwidth]{netPerfCoef_ameSR}
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

% tables
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{colortbl}

@@

% references to graphics
\makeatletter
\def\input@path{{/Users/janus829/Research/netsMatter/prezzy/graphics/}, {/Users/s7m/Research/netsMatter/prezzy/graphics/}}
\graphicspath{{/Users/janus829/Research/netsMatter/prezzy/graphics/}, {/Users/s7m/Research/netsMatter/prezzy/graphics/}}

\title[AMEN]{\textsc{Taking Dyads Seriously}}
\author[Minhas et al.]{Shahryar Minhas, Cassy Dorff, Max Gallop, Margaret Foster, Juan Tellez, Howard Liu, Michael Ward}

@@

\begin{itemize}
\item biased effects estimation
\item uncalibrated confidence intervals
\item poor predictive performance
\item inaccurate description of network phenomena
\end{itemize}

@@

\begin{tabular}{lll}
Thompson \& Walker (1982) & Beck et al. (1998) & Snijders (2011) \\
Frank \& Strauss (1986) & Signorino (1999) & Erikson et al. (2014) \\
Kenny (1996) & Li \& Loken (2002) & Aronow et al. (2015) \\
\end{tabular}

@@

\begin{itemize}
  \item Nodal and dyadic dependencies in networks
  \begin{itemize}
	  \item Can model using the ``A'' in AME 
  \end{itemize}
  \item Third order dependencies
  \begin{itemize}
	  \item Can model using the ``M'' in AME
  \end{itemize}	  
  \item Simulation	  
  \item Empirical Application
\end{itemize}

@@

We use this model to form the additive effects portion of AME (Warner et al. 1979, Li \& Loken 2002, Hoff 2005):

@@

Each node $i$ has an unknown latent factor:

@@

\vspace{-5mm}
The probability of a tie from $i$ to $j$ depends on their latent factors (Hoff 2008, Hoff 2015, Minhas, Hoff \& Ward 2018):

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{AME}

The full AME framework enables us to not only account for the types of dependencies that often arise in dyadic data but also estimate them:

\begin{align}
  \begin{aligned}
    y_{ij} \;=\; f(\theta_{ij}) &\text{, where } \\
    \theta_{ij} \;=\;& \bm\beta_{d}^{\top} \mathbf{X}_{ij} + \bm\beta_{s}^{\top} \mathbf{X}_{i} + \bm\beta_{r}^{\top} \mathbf{X}_{j} \text{\qquad(Exogenous parameters)} \\
    & + a_{i} + b_{j} + \epsilon_{ij} \text{\qquad\qquad\qquad\quad(SRRM parameters)} \\
    & + \mathbf{u}_{i}^{\top} \mathbf{D} \mathbf{v}_{j}  \text{\qquad\qquad\qquad\qquad\qquad\;(LFM parameters)} \\ 
  \label{eqn:ame}
  \end{aligned}
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Simulation}

Assume that the true data-generating process for a particular $Y$ is given by:

\vspace{-8mm}
\begin{align}
  y_{i,j} \sim  \mu + \beta x_{i,j} + \gamma w_{i,j} + \epsilon_{i,j}
  \label{eqn:sim}
\end{align}

We compare inference for $\mu$ and $\beta$---the latter parameter would be of primary concern for applied scholars---using three models:

\begin{itemize}
  \item the ``standard'' international relations approach estimated through a typical generalized linear model; 
  \item the AME approach outlined in the previous section with a unidimensional latent factor space ($K=1$);
  \item and an ``oracle'' regression model that assumes we have measured all sources of dependencies and thus includes both $x_{i,j}$ and $w_{i,j}$. 
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Bias Comparison}

We first compare the performance of the models in terms of how well they estimate the true values of $\mu$ and $\beta$:

\centering
\includegraphics[width=1\textwidth]{ameSimBias_all.pdf}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Coverage Comparison}

Next, we estimate the 95\% confidence interval for the three models in each of the simulations and estimate the proportion of times that the true value fell within those intervals:

\centering
\includegraphics[width=1\textwidth]{ameSimCover_all.pdf}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{What about the missing variable?}

Unobserved dependencies should be captured through the multiplicative effects portion of the model, $\mathbf{U}^{\top} \mathbf{D} \mathbf{V}$:

\centering
\includegraphics[width=1\textwidth]{ameSimCorr.pdf}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Empirical Applications}

Descriptive information about the replicated studies:

\footnotesize{
\begin{tabular}{lccccc}
	& Model &  Date Range & N. Actors  & Dyads Type & Clustering $\sigma_{\hat{\beta}}$ \\ \toprule
	Reiter \& Stam (2003) & Logit &1945--1995 &  193 & Directed & Robust \\ 
	Weeks (2012) & Logit & 1946--1999 & 197 & Directed & Robust \\
	Gibler (2017) & Logit & 1816--2008 & 193 & Undirected & Robust \\ \bottomrule
\end{tabular}
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Predictive Comparison}

\centering   
\begin{tabular}{lccc}
~ &  Model & AUC (ROC) & AUC (PR) \\ \toprule
\multirow{2}{*}{Reiter \& Stam (2003)} & \cellcolor{blue!10}AME & \cellcolor{blue!10}0.96 & \cellcolor{blue!10}0.15 \\
~ & GLM & 0.92 & 0.08 \\
\multirow{2}{*}{Weeks (2012)} & \cellcolor{blue!10}AME & \cellcolor{blue!10}0.97 & \cellcolor{blue!10}0.15 \\
~ & GLM & 0.64 & 0.00 \\
\multirow{2}{*}{Gibler (2017)} & \cellcolor{blue!10}AME & \cellcolor{blue!10}0.91 & \cellcolor{blue!10}0.08 \\
~ & GLM & 0.52 & 0.00 \\ \bottomrule
\end{tabular}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Results robust?}

\footnotesize{
\begin{table}[ht]
\centering
  \begin{tabular}{l p{4cm} l} \toprule
    \multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Confirmed after \\
    & &  accounting for dependencies? \\ \toprule
    Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Confirmed} \\ \midrule
    Weeks (2012) & Bosses, Juntas, and Strongmen are more Aggressive, Machines are Not & {Unconfirmed} \\\midrule
    Gibler (2017) & Power Parity at Time of Entry to International System Increases Conflict & {Unconfirmed}\\ \bottomrule
  \end{tabular}
\end{table}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{What else did we learn? Reiter \& Stam (2003)}

\begin{tabular}{cc}
  \includegraphics[width=.7\textwidth]{reiter_stam_aEff_map.pdf} &
  \includegraphics[width=.3\textwidth]{reiter_stam_aEff_line.pdf} \\
\end{tabular}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{What else did we learn? Weeks (2012)}

\centering
\includegraphics[width=.7\textwidth]{weeks_circPlot.png}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{What else did we learn? Gibler (2017)}

\centering
\includegraphics[width=\textwidth]{gibler_margeff.pdf}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \frame{
%   \frametitle{What's Next?}
%
%   \begin{itemize}
%     \item
%   \end{itemize}
%
% }
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Predictive Comparison: Reiter \& Stam (2003)}

\centering
\begin{tabular}{lr}
  \includegraphics[width=.45\textwidth]{reiter_stam_roc_outSample.pdf} & 
  \includegraphics[width=.45\textwidth]{reiter_stam_pr_outSample.pdf} \\
\end{tabular}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Predictive Comparison: Weeks (2012)}

\centering
\begin{tabular}{lr}
  \includegraphics[width=.45\textwidth]{weeks_roc_outSample.pdf} & 
  \includegraphics[width=.45\textwidth]{weeks_pr_outSample.pdf} \\
\end{tabular}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\frame{
  \frametitle{Predictive Comparison: Gibler (2017)}

\centering
\begin{tabular}{lr}
  \includegraphics[width=.45\textwidth]{gibler_roc_outSample.pdf} & 
  \includegraphics[width=.45\textwidth]{gibler_pr_outSample.pdf} \\
\end{tabular}

@@
